---
title: "DA R PROGRAMMING EXAM"
author: "DA class"
date: "22/7/2024"
output: html_document
---

 ## IMPORT THE DATASET CALLED employees_attrition.csv file 
 ## INTO RSTUDIO
 
 ## EMPLOYEE ATTRITION MEANS EMPLOYEES ARE LEAVING THE COMPANY
 ## FOR ONE REASON OR THE OTHER.
 
 ## THE ATTRITION COLUMN HAS TWO POSSIBLE VALUES YES - FOR EMPLOYEES
 ## THAT LEFT THE COMPANY AND NO FOR THOSE THAT ARE  STILL AROUND.
 
 ## NOTE
 
 ## 1. RUN EVERY CODE CHUNCK YOU CREATE SO THAT THE OUTPUT WILL
 ## BE DISPLAYED BELOW THE CODE CHUNK.
 
 ## 2. SAVE YOUR RMARKDOWN FILE WITH YOUR NAME FOR IDENTIFICATION
 
 ## 3. USE DPLYR FUNCTIONS TO ANSWER THE FOLLOWING QUESTIONS.
 
 
 
```{r Q1}

# load the tidyverse package into this session
library("tidyverse")
```
 
```{r  Q2}
## import the given employees_attrition.csv file 
identification <- read.csv(file.choose(), header = T)
```

```{r Q3 }

## SHOW THE STRUCTURE OF THE DATAFRAME
str(identification)
```


```{r Q4 }

## HOW MANY ROWS AND COLUMNS ARE THERE IN THE DATAFRAME
## There are 1470 rows and 17 columns in the data frame
```

```{r Q5}
## SHOW THE NUMBER OF EMPLOYEES THAT LEFT THE COMPANY AND THE 
## NUMBER THAT ARE STILL WITH THE COMPANY
## HINT - USE THE group_by() function to categorise.
identification %>%
  group_by(attrition) %>%
  summarise(employees = n())
```


```{r Q6}
## SHOW THE NUMBER OF EMPLOYEES THAT LEFT THE COMPANY IN EACH 
## DEPARTMENT.
identification %>%
  filter(identification$attrition == "Yes") %>%
  group_by(attrition, department) %>%
  summarise(employees = n())
```



```{r Q7}
## SHOW THE NUMBER OF MALE AND FEMALE EMPLOYEES THAT LEFT THE COMPANY
## AND THE NUMBER OF MALE AND FEMALE EMPLOYEES THAT ARE STILL WITH THE COMPANY
## HINT - USE THE group_by() function
identification %>%
  group_by(attrition, gender) %>%
  summarise(employees = n())
```




```{r Q8}

## WHAT IS THE AVERAGE NUMBER OF YEARS WORKED IN THE COMPANY BY THOSE THAT LEFT THE COMPANY.
identification %>%
  filter(identification$attrition == "Yes") %>%
  group_by(attrition) %>%
  summarise(employees = mean(TotalWorkingYears))
```



```{r Q9}

```




```{r Q10}
## SHOW THE NUMBER OF EMPLOYEES THAT LEFT THE COMPANY BASED ON  THE isMarried column. 
identification %>%
  filter(identification$attrition == "Yes") %>%
  group_by(attrition, isMarried) %>%
  summarise(employees = n())


```



```{r Q11 }
## SHOW THE AVERAGE MONTHLY INCOME EARNED BY THOSE WHO LEFT 
## THE COMPANY. BREAK THE INFORMATION DOWN BY DEPARTMENTS.
identification %>%
  filter(identification$attrition == "Yes") %>%
  group_by(attrition, department)  %>%
  summarise(income = mean(monthly_income))


```


```{r Q12}

## SHOW THE AVERAGE MONTHLY INCOME EARNED BY THOSE WHO ARE STILL WITH THE COMPANY. BREAK THE INFORMATION DOWN BY DEPARTMENTS.
identification %>%
  filter(identification$attrition == "No") %>%
  group_by(attrition, department)  %>%
  summarise(income = mean(monthly_income))
```



